---
title: "HD 5514 Research Methods (Fall 2019)"
subtitle: "Analysis of Variance (W15)"
author: "Koeun Choi"
date: "December 4, 2019"
output:
  rmdformats::readthedown:
    highlight: kate
    self_contained: false
    code_folding: hide
---

```{r knitr_init, echo=FALSE, cache=FALSE}
library(knitr)
library(rmdformats)
## Global options
options(max.print="75")
opts_chunk$set(echo=TRUE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

# In-Class Acitivity: Analysis of Variance (ANOVA)

## Load Data
We are going to use `R.appendix1.data` from Personality Project. We will load and print the data.
```{r}
# 1. Loading
data1 <- read.table("http://personality-project.org/r/datasets/R.appendix1.text", header = TRUE)   #read the data into a table
data1
# 2. Print
head(data1)
```

## Check Data Set
Check the number of rows and columns. 
```{r}
# Check the dimension of an object
dim(data1)

# Display the internal structure of an R object.
str(data1)

# Check the levels
levels(data1$Dosage)
```

## Visualize your data - Boxplot & Histogram
```{r}
#par(mar=c(1,1,1,1)) # change figure margins if you encounter error

# Boxplot
boxplot(data1$Alertness)    

# Boxplot (Alterness by group)
boxplot(data1$Alertness ~ data1$Dosage)

# Histogram
hist(data1$Alertness)    
```

## Conduct one-way Anova
```{r}
# Compute the analysis of variance
aov.ex1 <- aov(Alertness ~ Dosage, data = data1)
# Summary of the analysis
summary(aov.ex1)
```

## Report means
`model.tables` computes summary tables for model fits, especially  aov fits.
```{r}
# Computes summary table (mean, count)
model.tables(aov.ex1,"means")

?model.tables
```

## Check ANOVA assumptions
ANOVA assumes normal distribution and homogeneity of variances.

## Check the homogeneity of variance assumption: Levene’s test
Perform Leven's test using the function `leveneTest` in `car` package. If variance across groups is not significantly different, we can assume the homogeneity of variances in the different groups.
```{r}
library(car)
leveneTest(Alertness ~ Dosage, data = data1)
```

## ANOVA test with no assumption of equal variances: Welch one-way test
Welch one-way test is an alternative that does not require thae homogeneity of variance assumption.
```{r}
oneway.test(Alertness ~ Dosage, data = data1)
```

## Check the normality assumption
We can assume normality if all the points fall approximately along this reference line. Or you can conduct Shapiro-Wilk test on ANOVA residuals. 
```{r}
# create a normal QQ plot and add a reference line (Run the two lines at the same time)
qqnorm(cars$dist)
qqline(cars$dist) 

# Use plot function to get a QQ plot
plot(aov.ex1, 2)

# Run Shapiro-Wilk test
shapiro.test(x = residuals(aov.ex1))
```

## Non-parametric alternative to one-way ANOVA test: Kruskal-Wallis rank sum test
When ANOVA assumptions are not met, you cna use a non-parametric alternative.
```{r}
kruskal.test(Alertness ~ Dosage, data = data1)
```

# In-Class Acitivity: Multiple pairwise-comparison

## Bonferroni and Holm multiple pairwise-comparisons
In one-way Anova, a siginificant p-value means some of the group means are different. However, this does not tell us which pairs of groups are different. We can perform multiple pairwise-comparison to test the statistically significant mean difference between particular pairs of group. Given the significant ANOVA test, we can compute adjusted p-values using `pairwise.t.test`.
```{r}
# no adjustment
pairwise.t.test(data1$Alertness, data1$Dosage, data = , p.adj = "none")

# Bonferroni
pairwise.t.test(data1$Alertness, data1$Dosage, data = , p.adj = "bonf")

# Holm
pairwise.t.test(data1$Alertness, data1$Dosage, data = , p.adj = "holm")
```

## Tukey multiple pairwise-comparisons
Tukey HSD (Tukey Honest Significant Differences) using `TukeyHD`.
```{r}
# diff: difference between means of the two groups
#lwr, upr: the lower and the upper end point of the confidence interval at 95% (default)
# p adj: p-value after adjustment for the multiple comparisons.

TukeyHSD(aov.ex1)
```

# In-Class Acitivity: Planned Constrasts
You can test some specific hypotheses than testing all possible mean comparisons. We can combine multiple means from different levels and compare two means (e.g., a and b vs. c) 
```{r}
# assign values to the groups that you want to compare
c1 <- c(.5, -.5, 0) # H0_c1: a = b
c2 <- c(.5, .5, -1) # H0_c2: a & b = c

# assign the contrasts 
data1$c1 <- c(rep(.5,6), rep(-.5,8), rep(0,4))
data1$c2 <- c(rep(.5,6), rep(.5,8), rep(-1,4))

# compare differences between pairs of means (using regression)
anova(lm(Alertness ~ c1 + c2, data = data1))
```

# FYI: Visualize your data using ggpubr

## Bar plots with mean +/- se with jittered points
```{r}
# Install R package ggpubr

## Recommended: Install the latest developmental version from GitHub (remove # below)
#if(!require(devtools)) install.packages("devtools")
#devtools::install_github("kassambara/ggpubr")

## If that failed, try one from CRAN (Remove # below)
#install.packages("ggpubr")
#install.packages("ggplot2")

# Load ggpubr
library(ggpubr)
library(ggplot2)

ggbarplot(data1, x = "Dosage", y = "Alertness", add = c("mean_se", "jitter"))
```

# In-Class Acitivity: Tidyverse
`tidyverse`  is an opinionated collection of R packages designed for data science. https://www.tidyverse.org/ This is designed to make it easy to install and load core packages from the tidyverse.
`magrittr` offers a set of operators which make your code more readable such as the pipe operator `%>%` https://magrittr.tidyverse.org/
`dplyr` provides a set of tools for efficiently manipulating datasets: https://dplyr.tidyverse.org/.
```{r}
# The easiest way to get dplyr or magrittr is to install the whole tidyverse:
#install.packages("tidyverse")

# Alternatively, install just dplyr or magritter:
#install.packages("dplyr")
#install.packages("magrittr")

library(magrittr)
library(dplyr)
library(tidyverse)
```

## Computes summary tables using dplyr
```{r}
data1 %>% group_by(Dosage) %>%
  summarise(
    count = n(),
    mean = mean(Alertness, na.rm = TRUE),
    sd = sd(Alertness, na.rm = TRUE)
  )
```

# Assignment 10 (Week 15)

## Read Data
We will use  build-in data set `Prestige` contained in the R package `car`. We will load and print the survey data.
```{r}
# 1. Load the required package.
#install.packages("car") # if you don't have the "car" package, make sure to install it. 
library(car)

# 2. Load the data
data(Prestige) 

# 2. Print
head(Prestige)
```

## Check Data
Check the number of rows and columns. 
```{r}
# Display the internal structure of an R object.
str(Prestige)

# Check the dimension of an object
dim(Prestige)
```

## Use the help function to learn about variables
If you want to learn more about the t.test function.
```{r}
# A question mark is a shorcut for the "help" function.
?Prestige
```