---
title: "HD 5514 Research Methods (Fall 2019)"
subtitle: "Inference for Regression (W13)"
author: "Koeun Choi"
date: "November 20, 2019"
output:
  rmdformats::readthedown:
    highlight: kate
    self_contained: false
    code_folding: hide
---

```{r knitr_init, echo=FALSE, cache=FALSE}
library(knitr)
library(rmdformats)
## Global options
options(max.print="75")
opts_chunk$set(echo=TRUE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

# In-Class Acitivity: Correlation

## Load Data
We are going to use `cars` data included in R by default. We will load and print the cars data.
```{r, eval=FALSE}
# 1. Loading
data(cars) 

# 2. Print
head(cars)
```

## Check Data Set
Check the number of rows and columns. 
```{r, eval=FALSE}
# Display the internal structure of an R object.
str(cars)

# Check the dimension of an object
dim(cars)

# Check the variables
?cars
```

## Visualize your data using a scatter plot
Find the data columns, named `speed` (numeric	Speed (mph))  and `dist` (numeric	Stopping distance (ft)) from the cars data set. Create a scatter plot to visualze the relationship between the two variables.
```{r, eval=FALSE}
#par(mar=c(1,1,1,1)) # change figure margins if you encounter errors

# Create a simple scatter plot (Method 1)
plot(cars$dist ~ cars$speed) 

# Create a simple scatter plot (Method 2)
plot(dist ~ speed, data = cars) 
```

## Calcuate correlation
Calcuate correlation among two variables `speed` and `dist` using the `cor` function. Conduct statistical testing for the association using `cor.test` function. The default is Pearson's correlation, but you can use Spearman's (rho) rank correlation `method="spearman"` or Kendall's tau `method="kendall"` for nonparametic tests.
```{r, eval=FALSE}
# Compute correlation
cor(cars$speed, cars$dist)

# Statistical testing for associatin between the two variables
cor.test(cars$dist, cars$speed)

# Check the `cor.test` function
?cor.test 

# Correlations for the entire data set
cor(cars)
```

# In-Class Acitivity: Simple Linear Regression

## Load Data
We are going to use `cars` data included in R by default. We will load and print the cars data.
```{r, eval=FALSE}
# 1. Loading
data(cars) 

# 2. Print
head(cars)
```

## Check Data Set
Check the number of rows and columns. 
```{r, eval=FALSE}
# Display the internal structure of an R object.
str(cars)

# Check the dimension of an object
dim(cars)

# Check the variables
?cars
```

## Visualize your data using a scatter plot
Find the data columns, named `speed` (numeric	Speed (mph))  and `dist` (numeric	Stopping distance (ft)) from the cars data set. Create a scatter plot to visualze linear relationship between the two variables.
```{r, eval=FALSE}
#par(mar=c(1,1,1,1)) # change figure margins if you encounter errors

# Create a scatter plot
scatter.smooth(cars$dist ~ cars$speed) 

# Add axes and text
scatter.smooth(cars$dist ~ cars$speed,
main="Dist ~ Speed",
xlab="Speed (mph)",
ylab="Stopping distance (ft)",
col="orange")               
```

## Visualize your data using a boxplot and check for outliers
Use the `boxplot` funciton to create a boxplot for each variable: `speed` (numeric	Speed (mph))  and `dist` (numeric	Stopping distance (ft)) from the cars data set. See if there is any outlier based on the 1.5 x interquartile-range (1.5 * IQR) rule.
```{r, eval=FALSE}
# box plot for 'speed'
boxplot(cars$speed,  main="Speed")

# box plot for 'speed'
boxplot(cars$dist,  main="Distance")

# Visualize two figures side by side and add notes (run the following 3 rows at the same time)
par(mfrow=c(1, 2)) # divide graph area in 2 columns
boxplot(cars$speed, main="Speed", sub = paste("Outlier rows: ", boxplot.stats(cars$speed)$out))  
boxplot(cars$dist, main="Distance", sub=paste("Outlier rows: ", boxplot.stats(cars$dist)$out))  
```

## Check the normality of the response variable using a histogram
Create a histogram with the function `hist()`. You can change the number of bins using the `breaks=` argument.
```{r, eval=FALSE}
# simple histogram
hist(cars$dist)

# histogram with different number of bins
hist(cars$dist, breaks = 20)

# histogram with different number of bins and change color
hist(cars$dist, breaks = 20, col = "gray")
```

## Check the normality of the response variable using a density plot
Create a density plot with the function `density()`.
```{r, eval=FALSE}
# returns density data
density(cars$dist)

# plot the results (Method 1)
plot(density(cars$dist))

# plot the results (Method 2)
d <- density(cars$dist)
plot(d)
```

## Check the normality of the response variable using a QQ plot
Create a density plot with the function `qqnorm()`. Quantile-Quantile plot (QQ-plot) shows the correlation between a given sample and the normal distribution. 
```{r, eval=FALSE}
# create a normal QQ plot
qqnorm(cars$dist)

# create a normal QQ plot and add a reference line (Run the two lines at the same time)
qqnorm(cars$dist)
qqline(cars$dist) 
```

## Fit a simple linear regression model
Fit a linear model with the function `lm()`. Regress the outcome variable `dist` on the exploratory variable `speed`
```{r, eval=FALSE}
# fit the regression model (Method 1)
lm(cars$dist ~ cars$speed)

# fit the regression model (Method 2)
lm(dist ~ speed, data = cars)
```

## Produce a summary of the model fitting 
Produce a summary of the results of the model fitting using `summary` funciton.
```{r, eval=FALSE}
# check the summary
LinearModel <- lm(dist ~ speed, data = cars)
summary(LinearModel)
```

## Checking for statistical significance
Check coefficients, F-statistic, and R-squared
```{r, eval=FALSE}
LMSummary <- summary(LinearModel)

# model coefficients
LMSummary$coefficients

# model F-statistic
LMSummary$fstatistic
```

# In-Class Acitivity: Multiple Linear Regression

## Load Data
We are going to use `mtcars` data included in R by default. We will load and print the mtcars data.
```{r, eval=FALSE}
# 1. Loading
data(mtcars) 

# 2. Print
head(mtcars)
```

## Fit a multiple linear regression model
Fit a linear model with the function `lm()`. Regress the outcome variable `mpg`(Miles/(US) gallon) on the exploratory variables `cyl` (Number of cylinders), `wt` (Weight (1000 lbs)), and `vs` Engine (0 = V-shaped, 1 = straight).
```{r, eval=FALSE}
# fit the regression model
lm(mpg ~ cyl + wt + vs, data = mtcars)
?mtcars
```

## Produce a summary of the model fitting 
Produce a summary of the results of the model fitting using `summary` funciton.
```{r, eval=FALSE}
# check the summary
MulipleModel <- lm(mpg ~ cyl + wt + vs, data = mtcars)
summary(MulipleModel)
```

# Assignment 9 (Week 13)

## Read Data
We will use  build-in data set `Prestige` contained in the R package `car`. We will load and print the survey data.
```{r, eval=FALSE}
# 1. Load the required package.
#install.packages("car") # if you don't have the "car" package, make sure to install it. 
library(car)

# 2. Load the data
data(Prestige) 

# 2. Print
head(Prestige)
```

## Check Data
Check the number of rows and columns. 
```{r, eval=FALSE}
# Display the internal structure of an R object.
str(Prestige)

# Check the dimension of an object
dim(Prestige)
```

## Use the help function to learn about variables
If you want to learn more about the t.test function.
```{r, eval=FALSE}
# A question mark is a shorcut for the "help" function.
?Prestige
```

## Check three data frame columns (income, education)
First, find the data column, named `income`, of the survey data set.
```{r, eval=FALSE}
# Summary of the "income" column
summary(Prestige$income)
```

Next, find another data column, named `education` in the survey data set.  
```{r, eval=FALSE}
# Summary of the "group" column
summary(Prestige$education)
```

Lastly, find the data column, named `prestige`, of the survey data set.
```{r, eval=FALSE}
# Summary of the "prestige" column
summary(Prestige$prestige)
```

## Visualize your variable and check for outliers
Use the `boxplot` funciton to create a boxplot for each variable: `income` (Average income of incumbents, dollars, in 1971), `education` (Average education of occupational incumbents, years, in 1971), `prestige` (Pineo-Porter prestige score for occupation, from a social survey conducted in the mid-1960s) from the cars data set. See if there is any outlier based on the 1.5 x interquartile-range (1.5 * IQR) rule.
```{r, eval=FALSE}
?Prestige
# box plot for 'income'
boxplot(Prestige$income, main="Income")
boxplot(Prestige$income, main="Income", sub = paste("Outlier rows: ", boxplot.stats(Prestige$income)$out))  

# Check the outliers for 'income'
boxplot.stats(Prestige$income)$out

# box plot for 'education'
boxplot(Prestige$education, main="Education")

# Check the outliers for 'education'
boxplot(Prestige$education, main="Education", sub = paste("Outlier rows: ", boxplot.stats(Prestige$education)$out))  

# box plot for 'prestige'
boxplot(Prestige$prestige, main="Prestige")

# Check the outliers for 'prestige'
boxplot(Prestige$prestige, main="Prestige", sub = paste("Outlier rows: ", boxplot.stats(Prestige$prestige)$out))
```

## Check the normality of a potenital response variable using a histogram
Create a histogram with the function `hist()`. You can change the number of bins using the `breaks=` argument.
```{r, eval=FALSE}
# histogram with different number of bins and change color
hist(Prestige$income, breaks = 5, col = "gray")
```

## Check the normality of a potenital response variable using a density plot
Create a density plot with the function `density()`.
```{r, eval=FALSE}
# plot the results
d2 <- density(Prestige$income)
plot(d2)
```

## Check the normality of a potenital response variable using a QQ plot
Create a density plot with the function `qqnorm()`. Quantile-Quantile plot (QQ-plot) shows the correlation between a given sample and the normal distribution. 
```{r, eval=FALSE}
# create a normal QQ plot and add a reference line (Run the two lines at the same time)
qqnorm(Prestige$income)
qqline(Prestige$income) 
```

## Check the normality of the response variable using a histogram
Create a histogram with the function `hist()`. You can change the number of bins using the `breaks=` argument.
```{r, eval=FALSE}
# histogram with different number of bins and change color
hist(Prestige$prestige, breaks = 10, col = "gray")
```

## Chek the normality of the response variable using a density plot
Create a density plot with the function `density()`.
```{r, eval=FALSE}
# plot the results
d2 <- density(Prestige$prestige)
plot(d2)
```

## Check the normality of the response variable using a QQ plot
Create a density plot with the function `qqnorm()`. Quantile-Quantile plot (QQ-plot) shows the correlation between a given sample and the normal distribution. 
```{r, eval=FALSE}
# create a normal QQ plot and add a reference line (Run the two lines at the same time)
qqnorm(Prestige$prestige)
qqline(Prestige$prestige) 
```

## Fit a simple linear regression model
Fit a linear model with the function `lm()`. Regress the outcome variable `prestige` on the exploratory variable `education`
```{r, eval=FALSE}
# fit the regression model
lm(prestige ~ education, data = Prestige)
```

## Produce a summary of the model fitting 
Produce a summary of the results of the model fitting using `summary` funciton.
```{r, eval=FALSE}
# check the summary
LinearModel2 <- lm(prestige ~ education, data = Prestige)
summary(LinearModel2)
```